---
title: "Visualization_1"
author: "Kimberly Lopez"
date: "2024-09-26"
output: github_document
---
```{r}
library(tidyverse)
library(ggridges)

```


Importing weather data. 

- data is publicly avaialble
- rnoaa is allowing you to extract the data for the weathr stations & selectig data variables from 2021-2022
- mutating by creating a new variable "name" for case_matching and id and turning time into degrees celcious 
- keeping id and everything else the same 
- use df |> view() to view the df 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

```

# Basic Scatterplot 
Making Plot 1 using ggplot. 

- ggplot(dfname, aes(x = , y = ))
- this, alone, is a blank scatter plot becasue we have not stated the geometry yet
- use geom_point() to put the scatter plot points for this

```{r}
 ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

You can also do this as part of a collection of other codes like piping 

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y =tmax))+ 
  geom_point()

```

Saving ggplot objects by assigning it to a variable ggp_weather_scatterplot

```{r}
ggp_weather_scatterplot= 
  weather_df |> 
  ggplot(aes(x=tmin, y =tmax))+ 
  geom_point()
  
```

We are getting warming message of missing values. Use filter function to check which rows where there are missing data

```{r}
weather_df |>
  filter(is.na(tmax))
```

# Advanced scatterplot 

Use color in aes to assign the variables to a color
- alpha makes points more transparent in the geom_point changes are not specfiic to variables so it aplies to the whole plot 
- geom_smooth creates a smooth line in the points

**where you define the aesthetics matters**

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color= name)) + 
  geom_point(alpha = .3, size= .8) + 
  geom_smooth(se = FALSE)
```
Used aes in the scatterplot from the whole plot, so then it creates a smooth line to the whole plot 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```

Use faceting to have 3 seperate scatter plots that are not overlaying on each other. 
- left hand side is what variable we ave on the right hand side placing columns first 

```{r}

weather_df |>
  ggplot(aes(x= tmin, y = tmax, color = name)) + 
  geom_point (alpha = .3) + 
  geom_smooth (se= FALSE) + 
  facet_grid(. ~ name)
```

Plot using dates as the x-axis and y-amix to tmax to see seasonal trends in all 3 locations.
- setting size = prcp to show if it rained more on a certain day the dots will be bigger

```{r}

ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

Write a code chain that starts with weather_df

- focuses only on Central Park (use filter()) 
- converts temperatures to Fahrenheit (temp * (9/5)+32)
- makes a scatterplot of min vs. max temperature
- and overlays a linear regression line (using options in geom_smooth(method = "lm"))

```{r}
weather_df |> 
  filter( name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr= tmax * (9/5)+32, 
    tmin_fahr = tmin *(9/5)+32
  ) |>
ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()+ 
  geom_smooth(method = "lm", se=FALSE) 
  



```

Using geom_hex shows highest peaks of the data and data density. 

- setting the real color of the plot in geom_point() since there is not a color vairable in the data soo it cant go in the ggplot aes() row

```{r}
weather_df |>
  ggplot(aes(x=tmin, y = tmax))+ 
  geom_hex()

```
Using geom_hex shows highest peaks of the data and data density. 

- setting the real color of the plot in geom_point() since there is not a color vairable in the data soo it cant go in the ggplot aes() row

```{r}
weather_df |>
  ggplot(aes(x=tmin, y = tmax))+ 
  geom_hex()

```

# Univariable Plots (1 variable)

Use geom_histogram() to make a histogram of only variable tmax

```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()

```

Comparison of indiviaul places from the weather data next of each other. 

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

To fix this. Use facet_grid() function seperated by name of location. 


```{r}
weather_df |> 
  ggplot(aes(x=tmin, fill = name)) + 
  geom_histogram() + 
  facet_grid(.~name)
```

Instead we can also use a density plot like "smooth histogram: 
- this is the one professor would go for if comparing across locations

```{r}

weather_df |> 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_density (alpha = .3)
```


We can also do boxplots

```{r}

weather_df |>
  ggplot(aes(x=name, y = tmin, fill=name)) + 
  geom_boxplot()
```

Violin Plot: 

- used for vertical display of densities 
- **professor prefers violin plots compared to boxplots when the data is multimodal(shows up as more bumpy) or skewed**
- looking at distribution across tmin 

```{r}

weather_df |> 
  ggplot(aes(x=name, y = tmin, fill=name))+ 
  geom_violin()
```

Ridge Plots 

- used for separating each densities and separated vertically
- looks similar to density plot, separating each variable on the y axis
- must install ggridges


```{r}
weather_df |> 
  ggplot( aes(x = tmax, y = name)) + 
  geom_density_ridges()
```

Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable

- boxplot is the best since it shows the outliers the best 

```{r}
weather_df |> 
  ggplot( aes(y = prcp, x = name)) + 
  geom_boxplot() 
```

# Saving and Embedding plot


```{r}
ggp_weather = 
  ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 5)
```


