---
title: "Visualization_1"
author: "Kimberly Lopez"
date: "2024-09-26"
output: github_document
---
```{r}
library(tidyverse)

```


Importing weather data. 

- data is publicly avaialble
- rnoaa is allowing you to extract the data for the weathr stations & selectig data variables from 2021-2022
- mutating by creating a new variable "name" for case_matching and id and turning time into degrees celcious 
- keeping id and everything else the same 
- use df |> view() to view the df 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

```

# Basic Scatterplot 
Making Plot 1 using ggplot. 

- ggplot(dfname, aes(x = , y = ))
- this, alone, is a blank scatter plot becasue we have not stated the geometry yet
- use geom_point() to put the scatter plot points for this

```{r}
 ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

You can also do this as part of a collection of other codes like piping 

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y =tmax))+ 
  geom_point()

```

Saving ggplot objects by assigning it to a variable ggp_weather_scatterplot

```{r}
ggp_weather_scatterplot= 
  weather_df |> 
  ggplot(aes(x=tmin, y =tmax))+ 
  geom_point()
  
```

