---
title: "Visualization_2"
author: "Kimberly Lopez"
date: "2024-10-01"
output: github_document
---

```{r}
library(tidyverse)
library ( patchwork)
```

# Visualization 2

We’ll still work with NOAA weather data, which is loaded using the same code as in Vis. 1
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

# Labels 
Starting with scatterplot, but use labs function to label axis names

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Weather data from the rnoaa package for three stations"
  )
```

# Scales

Sometimes the default of ggplot scales this. Use scale_x_continous or scale_y_discrete to set breaks = to a number.
- you can transform the data in the scale_* function
- you can also set limits in this function 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15º C", "0", "15"),
     limits = c(-20, 30)) + 
   scale_y_continuous(
    trans = "sqrt", 
    position = "right")
  
```

# Color in ggplot

Use scale_color_hue to make different color pallets. Dont make your own color pallets ranges instead use set combinations viridis package 

- here we ad to tell the scale_color_viridis function that discrete = TRUE. 
- We used discrete = TRUE because the color aesthetic is mapped to a discrete variable.
- viridis::scale_fill_viridis() function is appropriate for the fill aesthetic used in histograms, density plots, and elsewhere.


```{r}
ggp_temp_plot= 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(h = c(100, 300))+ 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )
```


# Themes 

These control the overall appearance of the plots you are trying to make which is not specific to aesthtic mapping. Were only changing how the plots look in general. 

- legend.position = "none" will remove the legend. 
- theme_bw() will set background to white
- theme_minimal() is like the black and white 
- theme_classic () gives you the lines of the bars 

- order matters!!!!!
```{r}
ggp_temp_plot + 
  theme_bw() +
  theme(legend.position = "bottom")

```
```{r}
ggp_temp_plot + 
  theme_classic() + 
  theme(legend.position = "bottom")
```


Dont USe this:
```{r}
ggp_temp_plot + 
  ggthemes::theme_excel() + 
  theme(legend.position = "bottom")
```

 A Good plot example: 
 
Revisit the plot showing tmax against date for each location. Use labels, scale options, and theme changes to improve the readability of this plot
steps
- ggplot() 
- geom_smooth() standard error = false
- geom_point() to add the points 
- labs () labeling axis 


```{r}
weather_df|> 
  ggplot( aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) + 
  geom_point(aes(size = prcp), alpha = .75) + 
  labs(
    title = "Temperature plot",
    x = "Date",
    y = "Seasonal variation in Maxiumum temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE) + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```


